A modern, event-based architecture for distributed evolutionary algorithms

### 2018-04-03 paper.tex
References to BOBB and COCOM

  %%% Maybe a little more about the why Island?
  %%% Not only better performance in time, but also, more diverstity and less
  %%% parametrization
  % Please add something and I'll review it. - JJ
  % OK, TO DO - M
  % And I have done it now... - JJ

  %% TODO: 
  %% Explain a little more BBOB the benchmarks
  %% - Mario

  We will examine the results of this framework on some BBOB benchmarks,
  taking into account
  %speed as well as
  the number of evaluations, comparing it
  against another Cloud ready parallel pool based implementation (with
  containers).
  %The speed measurements are intended mainly as baseline,
  %since no effort has been made to optimize it.
  The
  implementation is also free software and can be downloaded from
  GitHub.
  %% TODO: Add link?

  The rest of the paper is organized as follows. Next we present the
  state of the art in cloud implementation of evolutionary algorithms,
  to be followed in section \ref{sec:methods} by an introduction to the
  serverless architecture we will be using as well as our mapping of the
  evolutionary algorithm to it. Section \ref{sec:res} will present the
  result of performing experiments with this proof of concept; finally
  in section \ref{sec:con} will discuss the results, present conclusions
  and future lines of work.

  \section{State of the art}

  In general, scientific computing has followed the trends of the
  computing industry, with new implementations published almost as soon
  as new technologies became commercially available, or even
  before. There were very early implementations of evolutionary
  algorithms on transputers \cite{voigt1990modelling}, the world wide
  % transputers, had to google that
  % you're too young, man - JJ
  web \cite{chong:1999:jDGPi} and the first generation of cloud
  services
  \cite{DBLP:journals/corr/abs-1105-6205,de2017parallel,salza2017ccube}. %Evolutionary
  %algorithms are population-based algorithms, so it is relatively easy
  %to create a parallel implementation by just splitting populations, as
  %long as there is a way for those populations to interchange some
  %individuals and leverage the use of many different machines. In
  %general, besides, one of the most time-consuming steps in an
  %evolutionary algorithm is evaluation, so as long as you can compute
  %fitness in parallel, you can obtain a certain amount of speed up.
  However, every new computing platform has its own view of computing, and in many cases that has made evolutionary algorithms move in new directions in order to make them work better in that platform while keeping the spirit of
  bio-inspiration. For instance, most evolutionary algorithms work in a
  synchronous way; although there were very early efforts to create
  asynchronous EAs \cite{coleman89}, in general
  generations proceed one after the other and migration happens in
  all islands at the same time. However, this mode of working does not
  fit well with architectures that are heterogeneous and dynamic, which
  is why there have been many efforts from early
  on to adapt EAs to this kind of substrate
  \cite{Jini:FEA2000,zorman2002creation,baugh2003asynchronous}.
  %This last one creates an asynchronous, parallel system that is {\em functionally equivalent} to a sequential one by {\em unrolling} operations so
  %that they happen in parallel as long as there is no dependence between
  %them; multiple generations occur at the same time with every operation
  %using data that is available at every particular moment.
  %\cite{Jini:FEA2000} and \cite{zorman2002creation} explicitly
  %follow two completely different, and opposed, approaches to the
  %adaptation of an EA to a heterogeneous substrate. While the former
  %tends towards finer grain by distributing every particular operation
  %to different Jini services, the latter uses a pool-based approach to
  %create island-like structures, creating again a system that, even
  %leveraging the capabilities of the architecture in a much better way,
  %does not advance the  structure of an EA in any new way.

@@

  The layers and message flow in the application are shown in Figure
  \ref{fig:kafkeo}, which also includes the evolutionary components. We
  will focus for the time being in the general picture: a serverless
  architecture using a messaging service as a backbone, which in this case
  takes the shape of Kafka/MessageHub. These messages are produced and
  consumed by a service, which can also store them in an external database
  for their later use; in general, messaging systems are configured to
  keep messages only for a certain amount of time, and they disappear
  after that. Messaging queues are organized in {\em topics} and every
  topic uses a series of {\em partitions}, which can be increased for
  bigger throughput; the
  functions, hosted in OpenWhisk, execute {\em actions} triggered by the
  arrival of new messages; these actions also produce new messages that
  go back to the MessageHub to continue with the message loop. If all
  this is hosted in a cloud provider, as it is the case, the MessageHub
  service will be charged according to a particular cost
  structure, with partitions taking the most part of
  the cost, while messages have a relatively small impact.

@@

  %
  % \begin{table}
  %   \small
  %   \caption{ DEAP GA EvoWorker Parameters }
  %   \label{tab:GAparams}
  %   \centering
  %   \small
  %   \begin{tabular}{|l|c|}
  %     \hline
  %     Selection & Tournament size=12\\ \hline
  %     Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05  \\ \hline
  %     Mutation Probability & [.1,.6]  \\ \hline
  %     Crossover & Two Point  \\ \hline
  %     Crossover Probability& [.8,1]  \\ \hline
  %   \end{tabular}
  % \end{table}
  %
  %
  \begin{figure*}[h!tb]
  \begin{tabular}{l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l}
  \includegraphics[width=0.266\textwidth]{ppfigdim_f001}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f006}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f002}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f007}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f003}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f008}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f004}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f009}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f005}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f010}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{\label{fig:aRTgraphs}
   \bbobppfigdimlegend{$f_1$ and $f_{24}$}
  }
  \end{figure*}
  %
  The results of the comparison are shown in Figure \ref{fig:aRTgraphs},
  which follow the classical BBOB 2009 format, which includes the amount
  of effort devoted to finding a certain fitness level and time needed
  to do it. The EvoSpace and KafkEO results are shown side by side, only
  for the sake of comparison, since we are only interested for the time
  being in the baseline performance of the proof of concept.

@@

  % On the implementation side, we will work on the automatic deployment
  % of this architecture to any free or vendor serverless implementation,
  % including maybe abstraction layers such as the {\tt serverless}
  % library.

>>>>

  We will examine the results of this framework using the first five
  functions of the Noiseless Black-Box-Optimization-Benchmarking (BBOB) testbed \cite{hansen2016coco} part of the COCO (COmparing Continuous Optimisers) platform for comparisons of real-parameter global optimisers \cite{hansen2016coco}. The framework is compared
  against another cloud-ready parallel pool based implementation.
  %The speed measurements are intended mainly as baseline,
  %since no effort has been made to optimize it.
  The
  implementation is also free software and can be downloaded from
  GitHub.
  %% TODO: Add link? 
  The rest of the paper is organized as follows. Next we present the
  state of the art in cloud implementation of evolutionary algorithms,
  to be followed in section \ref{sec:methods} by an introduction to the
  serverless architecture we will be using as well as our mapping of the
  evolutionary algorithm to it. Section \ref{sec:res} will present the
  result of performing experiments with this proof of concept; finally
  in section \ref{sec:con} will discuss the results, present conclusions
  and future lines of work.

  \section{State of the art}
  In general, scientific computing has followed the trends of the
  computing industry, with new implementations published almost as soon
  as new technologies became commercially available, or even
  before. There were very early implementations of evolutionary
  algorithms on transputers \cite{voigt1990modelling}, the world wide
  % transputers, had to google that
  % you're too young, man - JJ
  web \cite{chong:1999:jDGPi} and the first generation of cloud
  services
  \cite{DBLP:journals/corr/abs-1105-6205,de2017parallel,salza2017ccube}. %Evolutionary
  %algorithms are population-based algorithms, so it is relatively easy
  %to create a parallel implementation by just splitting populations, as
  %long as there is a way for those populations to interchange some
  %individuals and leverage the use of many different machines. In
  %general, besides, one of the most time-consuming steps in an
  %evolutionary algorithm is evaluation, so as long as you can compute
  %fitness in parallel, you can obtain a certain amount of speed up.
  However, every new computing platform has its own view of computing, and in many cases that has made evolutionary algorithms move in new directions in order to make them work better in that platform while keeping the spirit of
  bio-inspiration. For instance, most evolutionary algorithms work in a
  synchronous way; although there were very early efforts to create
  asynchronous EAs \cite{coleman89}, in general
  generations proceed one after the other and migration happens in
  all islands at the same time. However, this mode of working does not
  fit well with architectures that are heterogeneous and dynamic, which
  is why there have been many efforts from early
  on to adapt EAs to this kind of substrate
  \cite{Jini:FEA2000,zorman2002creation,baugh2003asynchronous}.
  %This last one creates an asynchronous, parallel system that is {\em functionally equivalent} to a sequential one by {\em unrolling} operations so
  %that they happen in parallel as long as there is no dependence between
  %them; multiple generations occur at the same time with every operation
  %using data that is available at every particular moment.
  %\cite{Jini:FEA2000} and \cite{zorman2002creation} explicitly
  %follow two completely different, and opposed, approaches to the
  %adaptation of an EA to a heterogeneous substrate. While the former
  %tends towards finer grain by distributing every particular operation
  %to different Jini services, the latter uses a pool-based approach to
  %create island-like structures, creating again a system that, even
  %leveraging the capabilities of the architecture in a much better way,
  %does not advance the  structure of an EA in any new way.

@@

  The layers and message flow in the application are shown in Figure
  \ref{fig:kafkeo}, which also includes the evolutionary components. We
  will focus for the time being in the general picture: a serverless
  architecture using a messaging service as a backbone, which in this case
  takes the shape of the Kafka/MessageHub service. 
  These messages are produced and
  consumed by a service, which can also store them in an external database
  for their later use; in general, messaging systems are configured to
  keep messages only for a certain amount of time, and they disappear
  after that. Messaging queues are organized in {\em topics} and every
  topic uses a series of {\em partitions}, which can be increased for
  bigger throughput; the
  functions, hosted in OpenWhisk, execute {\em actions} triggered by the
  arrival of new messages; these actions also produce new messages that
  go back to the MessageHub to continue with the message loop. If all
  this is hosted in a cloud provider, as it is the case, the MessageHub
  service will be charged according to a particular cost
  structure, with partitions taking the most part of
  the cost, while messages have a relatively small impact.

@@

  \begin{figure*}[h!tb]
  \begin{tabular}{l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l}
  \includegraphics[width=0.266\textwidth]{ppfigdim_f001}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f006}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f002}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f007}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f003}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f008}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f004}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f009}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f005}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f010}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{\label{fig:aRTgraphs}
   \bbobppfigdimlegend{$f_1$ and $f_{24}$}
  }
  \end{figure*}
  %
  The results of the comparison are shown in Figure \ref{fig:aRTgraphs},
  which follow the classical BBOB 2009 format, which includes the amount
  of effort devoted to finding a certain fitness level and time needed
  to do it. The EvoSpace and KafkEO results are shown side by side, only
  for the sake of comparison, since we are only interested for the time
  being in the baseline performance of the proof of concept.

@@
